Using HMMs for pre-snap coverage detection

### 2024-11-22 submission.Rmd
.

Choosing between man- and zone coverage is one of the most important strategic decisions of a defensive coordinator prior to each play in American football. While experienced coaches and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to infer the underlying tactic.
This project aims to leverage hidden Markov models (HMMs) to detect defensive strategies --- man or zone coverage --- based on pre-snap player movement data. 
By modeling hidden states that represent the offensive player to be guarded, this approach provides a data-driven framework for unraveling the complexity of defensive patterns and enables real-time tactical insights for coaching and analysis.

@@

#### 1. Pre-Motion Analysis
We train a model to predict whether the defense plays a man- or zone coverage scheme. The 
model uses the previously described features, blablabla and as observations information from those plays without pre-snap motion and information until the first motion for plays with motion.

#### 2. HMM Analysis
We model the movements of defensive players during the phase of pre-snap motion within a hidden Markov framework, in which the underlying states represent the offensive players to be guarded (see Franks et al., 2015 for a similar approach in basketball). In contrast to Groom et al., 2024, who enforce a state to proxy zone coverage during corner kicks in soccer, we cannot proceed similarly as the classical coverage zones in American football will only be covered post-snap. To remediate this, we derive the decision of man- or zone coverage from the number of switches for individual players. In particular, a low number of switches when offensive players are in motion indicates man coverage whereas a higher number indicates zone coverage.

#### 3. Post-Motion Analysis
We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, however, in this step, we incorporate the results from the HMM analysis.

>>>>

Choosing between man- and zone coverage is one of the most important strategic decisions a defensive coordinator has to take prior to each play in American football. While experienced coaches and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to infer the underlying tactic.
This project aims to leverage hidden Markov models (HMMs) to detect defensive strategies --- man or zone coverage --- based on pre-snap player movement data. 
By modeling hidden states that represent the offensive player to be guarded, this approach provides a data-driven framework for unraveling the complexity of defensive patterns and enables real-time tactical insights for coaching and analysis.

@@

#### 1. Pre-motion analysis
We train a model to predict whether the defense plays a man- or zone coverage scheme. The 
model uses the previously described features, blablabla and as observations information from those plays without pre-snap motion and information until the first motion for plays with motion.

#### 2. HMM analysis
We model the movements of defensive players during the phase of pre-snap motion within a hidden Markov framework, in which the underlying states represent the offensive players to be guarded (see Franks et al., 2015 for a similar approach in basketball). In contrast to Groom et al., 2024, who enforce a state to proxy zone coverage during corner kicks in soccer, we cannot proceed similarly as the classical coverage zones in American football will only be covered post-snap. To remediate this, we derive the decision of man- or zone coverage from the number of switches for individual players. In particular, a low number of switches when offensive players are in motion indicates man coverage whereas a higher number indicates zone coverage.

#### 3. Post-motion analysis
We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, however, in this step, we incorporate the results from the HMM analysis.
